---
title: 'Stats 101A Homework #4'
author: "Damien Ha"
date: "2023-02-24"
output: 
  pdf_document:
    toc: true
    toc_depth: 3
---


# Problem 1
## Loading Data

```{r}
papercompany <- read.table("papercompany.txt", header = T)
```


## Part A
```{r}
cor(papercompany)
```
\pagebreak
```{r}
pairs(papercompany, main = "Scatterplot Matrix of papercompany data", col = "turquoise", pch = 20)
```

All pairings of variables seem to have a strong, positive linear relationship with correlation coefficients close to positive 1. Certain pairs appear to have some odd trends or possible outliers, such as the points branching off and forming a curve in labor vs overhead. Overall each variable seems to have a positive trend.

## Part B
```{r}
model <- lm(papercompany$COST ~ papercompany$PAPER + papercompany$MACHINE +  
             papercompany$OVERHEAD + papercompany$LABOR)
summary(model)
```

The regression equation is $Y = 51.72314 + 0.94794X_1 + 2.47104X_2 + 0.04834X_3 - 0.05058X_4$ where $Y$ is cost and $X_1, X_2, X_3,$ and $X_4$ are paper, machine, overhead, and labor respectively

## Part C
```{r}
anova(model)
```

Paper has F = 18388.2129 with p = $2.2 \cdot 10^{-16}$ and machine has F = 126.8547 with p = $1.33 \cdot 10^{-10}$ suggesting that at $\alpha = 0.05$, these variables help better fit the model to the data. However, overhead has a p-value if 0.8711 and labor has a p-value of 0.2226, which, if we consider at the level $\alpha = 0.05$, may suggest these variables aren't significant

## Part D
From out linear model summary, we can see the $R^2$ value is 0.998, so 99.8% of the variation in cost is explained by the regression model  

## Part E
```{r}
# R-squared for
# cost vs paper
summary(lm(papercompany$COST ~ papercompany$PAPER))[[9]]
summary(lm(papercompany$COST ~ papercompany$PAPER + papercompany$MACHINE +   
             papercompany$OVERHEAD + papercompany$LABOR))[[4]][[2]]

# cost vs machine
summary(lm(papercompany$COST ~ papercompany$MACHINE))[[9]]
summary(lm(papercompany$COST ~ papercompany$PAPER + papercompany$MACHINE +   
             papercompany$OVERHEAD + papercompany$LABOR))[[4]][[3]]

# cost vs overhead
summary(lm(papercompany$COST ~ papercompany$OVERHEAD))[[9]]
summary(lm(papercompany$COST ~ papercompany$PAPER + papercompany$MACHINE +   
             papercompany$OVERHEAD + papercompany$LABOR))[[4]][[4]]

# cost vs labor
summary(lm(papercompany$COST ~ papercompany$LABOR))[[9]]
summary(lm(papercompany$COST ~ papercompany$PAPER + papercompany$MACHINE +   
             papercompany$OVERHEAD + papercompany$LABOR))[[4]][[5]]
```


## Part F
```{r}
reduced <- lm(papercompany$COST ~ papercompany$PAPER + papercompany$MACHINE)
anova(reduced, model)
```
F = 0.8 and p = 0.4615 so at $\alpha = 0.05$ we fail to reject the null hypothesis and say overhead and labor don't contribute significant information to cost when considering paper and machine variables

# Question 2
## Loading Data
```{r}
latour <- read.table("Latour.txt", header = T)
```


## Part A
```{r}
summary(lm(Quality ~ EndofHarvest + Rain + Rain:EndofHarvest, data = latour))
```
The interaction term has a p-value of 0.012 which, at $\alpha = 0.05$ is significant, suggesting the rate of change in quality rating depends on whether there has been any unwanted rain at vintage.

## Part B

### Part I

From the equation, letting $Y$ = quality, $X_1$ =  end of harvest days, and $X_2$ = rain, we have $Y = 5.1622 - 0.03145 X_1 + 1.7867 X_2 - 0.08314 X_1 X_2$ 

Setting $X_2$ = 0, we have $Y = 5.16122 - 0.03145 X_1$ 

The derivative of the equation with respect to $X_1$ is -0.03145, and we want to find a quality drop of 1:
```{r}
paste0("The number of days for a decrease in quality by 1 without rain is ", -1/-0.03145)
```

### Part II

Set $X_2$ = 1

$Y = 6.94792 - 0.11459 X_1$

The derivative with respect to $X_1$ is -0.11459, and we want to find a quality drop of 1
```{r}
paste0("The number of days for a decrease in quality by 1 with rain is ", -1/-0.11459)
```



